A meta-analysis of powdery mildew yield loss mitigation strategies in Australian mungbean

### 2020-07-13 08_MetaAnalysis_Emerson.Rmd
Corrected meta-analysis by Emerson based on the arm-based approach

@@

```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   filter(fungicide_ai != "control") %>% #remove controls from the data
   mutate(spray_management =
             factor(
                spray_management,
                c(
                   "Early",
                   "Recommended",
                   "Recommended_plus",
                   "Late",
                   "Late_plus"
                )
             ),
          D_pres = 
             factor(D_pres,
                    c("lowD",
                      "highD")),
          grain_MD = grain_yield.t.ha - yi_C) %>%
   select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype ,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      grain_SMD,
      grain_MD,
      dose,
      vi,
      vi2
      )
```

@@

```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management - 1,
   method = "ML",
   random = ~ spray_management | trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```
```{r}
(PM_mv$b - PM_mv$ci.ub) / PM_mv$se
```

In this result we can see that the `Early` treatment is not significantly different from zero, which in this case is the mean of the no spray `control`.
However the other treatments are significantly different from zero.
The $Q_M$ [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, shows the moderators significantly influence the model ($Q_M = 26.22, df = 5, p < 0.0001$) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010].
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 568.31, df = 111, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

Lets look at the trial variability in this model with a forest plot
```{r forestP, fig.height= 40, fig.width=17}
source(here("R/change_labels.R"))
source(here("R/forest_rows.R"))

### Sort and arrange data-frame occordingly and produce a index vector to inform the order of each "Trial" on the plot
index1 <- dat1 %>%
   mutate(ID = seq_along(trial_ref)) %>%
   mutate(spray_management = 
             factor(spray_management, 
                    levels = c("Early","Late","Recommended",
                    "Recommended_plus","Late_plus"))) %>%
   arrange(desc(spray_management), location, desc(grain_MD)) %>%
   mutate(ID2 = order(ID))%>%
   pull(ID2)

### Use forest rows function to return a list of where to place on the plot: sub-titles, each trial row, and summary prediction for each treatment.
### this list can be used below to reference where to plot each variable
rows1 <- forest_rows(head_row = 156, 
                     ord_var = dat1$spray_management, 
                     index = index1, 
                     gap = 5, 
                     row_offset = -2)

### A data-frame of meta-model coefficients used to produce the summary polygons for each treatment
polys <- data.frame(ref = change_labels(names(coef(PM_mv))),
                    x = coef(PM_mv),
                    ci.lb = PM_mv$ci.lb,
                    ci.ub = PM_mv$ci.ub)[c(1,4,2,3,5),]
polys$rows <- rows1$pred_row - 1.5

forest1  <- metafor::forest(PM_mv, cex = 1.15 , width = 5,
                #order = "fit",
                rows = rows1$trows-1.5,
                ilab = cbind(dat1$year, 
                             dat1$row_spacing, dat1$host_genotype, dat1$dose),
                ilab.xpos = c(-2.35,-2.1,-1.8,-1.4),
                slab = NA, 
                xlim = c(-2.9, 2.2),
                ylim = c(18,157),
                xlab = "Grain yield mean difference",
                addfit = FALSE,
                efac = c(0.2,0.15,0.2),
                at = seq(-1,1.5, 0.5))

### code which puts lines and annotations on the plot, its a bit long...
source(here("R/add_annotations.R"))

@@

Let's see if we can improve this model by including some other random effects and improving the random effect structure in the model.

```{r rma_noRandomInteractiveTerm}
PM_mv_St <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ 1 | trial),
   data = dat1
)

PM_mv_St1 <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(
      ~ 1 |
         location / year / trial_ref / dose / host_genotype / factor(row_spacing) /
         spray_management
   ),
   data = dat1
)

AIC(PM_mv) # Printing the AIC of the model helps differentiate the models in the anova output
anova(PM_mv, PM_mv_St) # PM_mv is the better model
AIC(PM_mv_St)
anova(PM_mv_St, PM_mv_St1) # PM_mv_St is the better model
```

The simpler `PM_mv_St` model explains significantly less variation than the more complicated model `PM_mv`.
So we will keep the original model `PM_mv`.

I have also tested the structure for testing the random trial and spray_management variables, these show that there is no significant difference to the models `PM_mv_St` and `PM_mv_St1` and that the originally method of specifying the random variables within the `trial` factor provides us with the best model.

The first table in this output shows the tau^2 (variance) of each random effects and the number of occurrences for each treatment in the analysis.
The second table is in two parts (left and right). 
The left part, rho, is the correlation of variation between the specified treatments.
All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho.
`Early` and `Late_plus` treatments never occurred within the same trial, which is indicated by the right side of the table (hence the warning).

The fixed effects, in the last table, show that yields in single early spray treatments are not significantly different from zero, or the no spray control.

@@

### Metafor comparisons

To make it easier to compare each of the treatments we can compute the meta-analysis contrasts.

```{r metafor_contrasts}
meta_cont <- anova(PM_mv, L = rbind(
   c(-1, 1, 0, 0, 0),
   c(-1, 0, 1, 0, 0),
   c(-1, 0, 0, 1, 0),
   c(-1, 0, 0, 0, 1),
   c(0, -1, 1, 0, 0),
   c(0, -1, 0, 1, 0),
   c(0, -1, 0, 0, 1),
   c(0, 0, -1, 1, 0),
   c(0, 0, -1, 0, 1),
   c(0, 0, 0, -1, 1)
))

change_labels <- function(x1) {
   x1 <- gsub(pattern = "spray_management",
              replacement = "",
              x = x1)
   x1
}

source("R/p_star.R")

table3 <- data.frame(
   contrasts = change_labels(meta_cont$hyp[, 1]),
   estimates = meta_cont$Lb,
   se = meta_cont$se,
   z_val = meta_cont$zval,
   pval = meta_cont$pval,
   VisP = p_star(meta_cont$pval)
)
write.csv(table3, here("cache/Table3_contrasts.csv"), row.names = FALSE)
```

@@

efficacy <- as_tibble(results_AI)
efficacy$Treatment <-
   factor(c(
      "Early",
      "Recommended",
      "Recommended_plus",
      "Late",
      "Late_plus"
   ))

@@

```{r metafor_plot}
efficacy %>%
   mutate(Treatment = factor(Treatment, levels = rev(
      c(
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))) %>%
   ggplot(aes(Treatment, Mean)) +
   geom_hline(
      yintercept = seq(-0.05, 0.3, by = 0.05),
      color = usq_cols("usq charcoal"),
      linetype = 3
   ) +
   geom_hline(yintercept = 0) +
   geom_point(aes(size = 1 / SE), shape = 15) +
   geom_linerange(aes(ymin = CIs_lower, ymax = CI_upper)) +
   coord_flip()
```

### Disease pressure effect

We also want to know if disease pressure exerts an effect on the decision for spray control.
Above we can see spraying at the first sign of powdery mildew or with in `r min(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"| slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` - `r max(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"|                   slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` days after first sign is effective at mitigating yield loss.
Does this change between years of low and high disease pressure.

We will add the intercept back into this model for ease of interpretation. 
```{r D_pres-Meta}
PM_mv_Dp <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management * D_pres - 1,
   method = "ML",
   random = ~ spray_management | trial,
   struct = "UN",
   data = dat1
)

summary(PM_mv_Dp)
```

These results show that the spray management variable we tested are not significantly better or worse between the two levels of disease pressure.
It seems that disease preasure does not improve the models ability to assess the most effective fungicide spray management option. Let's compare the disease pressure model to the original model to test this assumption.

```{r compareDvNDP}
AIC(PM_mv_Dp)
anova(PM_mv_Dp, btt = "D_pres")
anova(PM_mv, PM_mv_Dp)
```

A Wald test of `D_pres` shows the moderator is not significant in accounting for heterogeneity (P = 0.4859).
No significant difference which informs us to use the original model `PM_mv`, which is the simplest.

@@

```{r netmeta-analysis}
datPM3 <- slimmer_PM_dat %>%
   group_by(trial, spray_management, n) %>%
   summarize(yi_mean = mean(grain_SMD),
             vi_mean = mean(yield_error)) %>%
   ungroup()

@@

```{r netmeta-forest}
forest(
   net_con,
   reference.group = 5,
   rightcols = c("effect", "ci", "Pscore"),
   rightlabs = "P-Score",
   small.values = "bad"
)
```

>>>>

---
output: html_document
editor_options: 
  chunk_output_type: inline
---

@@

```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   mutate(vi = vi/n) %>% 
   #filter(fungicide_ai != "control") %>% #remove controls from the data
   mutate(spray_management =
             factor(
                spray_management,
                c(
                   "control",
                   "Early",
                   "Recommended",
                   "Recommended_plus",
                   "Late",
                   "Late_plus"
                )
             ),
          D_pres = 
             factor(D_pres,
                    c("lowD",
                      "highD")),
          grain_MD = grain_yield.t.ha - yi_C) %>%
   select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype ,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      grain_SMD,
      grain_MD,
      dose,
      vi,
      yield_error,
      vi2,
      n
      )
```

@@

```{r Metafor-analysis}
dat1$id <- 1:nrow(dat1)

PM_mv <- rma.mv(
   yi = grain_yield.t.ha,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(~ spray_management | trial, ~1 | id), 
   struct = "UN",
   control=list(optimizer="optim"),
   data = dat1
)

summary(PM_mv)
library(multcomp)
summary(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))), test=adjusted("none"))
plot(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))))

PM_mv2 <- rma.mv(
   yi = grain_yield.t.ha,
   vi,
   mods = ~ spray_management*D_pres,
   method = "ML",
   random = list(~ spray_management | trial, ~1 | id), 
   struct = "UN",
   control=list(optimizer="optim"),
   data = dat1
)
summary(glht(PM_mv2, linfct=cbind(contrMat(rep(7,12), type="Tukey"))), test=adjusted("none"))
plot(glht(PM_mv2, linfct=cbind(contrMat(rep(7,12), type="Tukey"))))

anova(PM_mv, PM_mv2)

@@

@@

@@

efficacy <- as_tibble(results_AI)
efficacy$Treatment <-
   factor(c(

       "control",
      "Early",
      "Recommended",
      "Recommended_plus",
      "Late",
      "Late_plus"
   ))

@@

```{r metafor_plot}
efficacy %>%
   filter(Treatment != "control") %>% 
   mutate(Treatment = factor(Treatment, levels = rev(
      c(

         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))) %>%
   ggplot(aes(Treatment, Mean)) +
   geom_hline(
      yintercept = seq(-0.05, 0.3, by = 0.05),
      color = usq_cols("usq charcoal"),
      linetype = 3
   ) +
   geom_hline(yintercept = 0) +
   geom_point(aes(size = 1 / SE), shape = 15) +
   geom_linerange(aes(ymin = CIs_lower, ymax = CI_upper)) +
   coord_flip()
```

@@

```{r netmeta-analysis}
datPM3 <- slimmer_PM_dat %>%
   group_by(trial, spray_management, n) %>%
   summarize(yi_mean = mean(grain_yield.t.ha),
             vi_mean = mean(yield_error)) %>%
   ungroup()

@@

```{r netmeta-forest}
forest(
   net_con,
   reference.group = 1,
   rightcols = c("effect", "ci", "Pscore"),
   rightlabs = "P-Score",
   small.values = "bad"
)
```
